---
title: "Dataset Exploration"
subtitle: "Exploring Customer Sales Data"
author: "Matt McGehee (Advisor: Dr. Seals)"
date: last-modified
date-format: long
format:
  html:
    code-fold: true
    embed-resources: true
    self-contained-math: true
course: Capstone Projects in Data Science
bibliography: references.bib # file contains bibtex for references
#always_allow_html: true # this allows to get PDF with HTML features
execute: 
  warning: false
  message: false
editor: 
  markdown: 
    wrap: 72
---

The dataset contains sales transaction details from a B2B context. The data tracks information on sales orders, quantities ordered, shipping details, and item-level statuses. Below is the data definition table that explains the key attributes.

```{r}
data_def <- data.frame(
  Attribute = c("OPCO", "SalesOrderID", "CustomerPO", "Product", "Description", 
                "Substrate", "RequestedDeliveryDate", "DateFulfilled", 
                "qtyOrdered", "qtyFulfilled", "UnitPrice", "TotalPrice", 
                "Class", "ShipToName", "ShipToAddress", "SalesOrderStatus", "SalesItemStatus"),
  Format = c("Varchar", "Numeric", "Numeric", "Varchar", "Varchar", 
             "Varchar", "Date", "Date", 
             "Numeric", "Numeric", "Numeric", "Numeric", 
             "Varchar", "Varchar", "Varchar", "Varchar", "Varchar"),
  Description = c("The customer placing the order, typically a distributor.",
                  "Unique identifier assigned to each sales order.",
                  "Customer's identifier for their order, sent to BCC.",
                  "Unique identifier assigned to each product.",
                  "Description of the product being sold.",
                  "Type of product/material.",
                  "Date the delivery was scheduled originally.",
                  "Date the delivery was made.",
                  "Quantity ordered on the order.",
                  "Quantity delivered on the order.",
                  "Price of each case of product SSI charges the customer.",
                  "Total price of the sales order.",
                  "Customer name.",
                  "Name of the ordering party.",
                  "Address where the order needs to be delivered.",
                  "Status of sales order.",
                  "Status of each line item on the sales order.")
)

knitr::kable(data_def, caption = "Sales Transaction Dataset Definition", align = "l")
```

### Sample Data
```{r}
sales_data <- read.csv("data_as_csv.csv")
DT::datatable(head(sales_data), options = list(pageLength = 6), 
          caption = "Interactive Preview of Sales Data")
```

### Summary Stats
```{r}
summary(sales_data)
```

This dataset allows for various analyses including:

-   Order fulfillment performance: By comparing the qtyOrdered and qtyFulfilled attributes, and the RequestedDeliveryDate against the DateFulfilled, one can analyze how efficiently orders are being fulfilled and whether delivery deadlines are met.

-   Sales trends and pricing: The UnitPrice and TotalPrice fields, alongside Product and Class, provide opportunities to evaluate product pricing strategies and customer buying patterns.

-   Customer segmentation: With attributes like Class and ShipToName, you can analyze customer behavior and segment them based on order sizes, frequency, or geography.

This dataset provides a detailed overview of the company's sales transactions. By leveraging these attributes, one can assess order performance, customer segmentation, and pricing strategies.
